### 2016-09-14 paper.tex
lots of changes

\documentclass[twoside]{article}
\usepackage{aistats2015}

@@

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{xcolor}
\usepackage[inline]{enumitem}
%\usepackage[authordate,bibencoding=auto,strict,backend=biber,natbib]{biblatex-chicago}
\usepackage[round]{natbib}   % omit 'round' option if you prefer square brackets
\bibliographystyle{plainnat}

@@

\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\W}{\mathcal{\hat W}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\wreopt}{\hat\w^{reopt}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\hat\w^{ave,*}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

\newcommand{\I}{I}
\newcommand{\Iinv}{I^{-1}}
\newcommand{\law}{\ensuremath{\xrightarrow{L}}}
\newcommand{\normal}[2]{\ensuremath{\mathcal{N}\left({{#1}},{{#2}}\right)}}
\newcommand{\trans}[1]{\ensuremath{{#1}^{\mathsf{T}}}}
\newcommand{\ltwo}[1]{{\left\lVert {#1} \right\rVert}_2}
\newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
\newcommand{\lzero}[1]{{\lVert {#1} \rVert}_0}
\newcommand{\proj}[1]{\pi_{{#1}}}
\newcommand{\prob}[1]{\Pr\left[{#1}\right]}

@@

\section{INTRODUCTION}
We consider the problem of distributed Maximum Likelihood Estimation (MLE).

%In a Generalized Linear Model,
%\begin{equation}
%\begin{aligned}
%y &= g^{-1}(\trans\x\w) + \epsilon
%,~~~
%\epsilon \sim \normal{0}{\sigma}
%\end{aligned}
%\end{equation}

@@

Let $\Y\subseteq\mathbb{R}$ be the space of response variables,
$X\subseteq\mathbb{R}^d$ be the space of the covariates,
and $\Theta\subseteq\mathbb{R}^d$ be the parameter space.
We assume a linear model.
The log-likelihood of data point $(\x,y)\in\X\times\Y$ given the model's true parameter $\wstar\in\Theta$ is denoted by $f(y,\trans\x\wstar)$.
%\begin{equation}
%y = g^{-1}(\trans\x\wstar)
%\end{equation}
%Here, $g : \mathbb{R} \to \mathbb{R}$ is the link function and $\wstar\in\Theta\subseteq\mathbb{R}^d$ is the true parameter of our model.
Let $Z\subset\X\times\Y$ be a dataset of $mn$ i.i.d. observations.
Then the maximum likelihood estimator is
\begin{equation}
%\wmle=\argmax_\w \sum_{i=1}^{nm} g(y_i-\trans\x_i\w)
\wmle=\argmax_\w \sum_{(\x,y)\in Z} f(y,\trans\x\w)
\end{equation}
Assume that $Z$ has been split up onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
Then each machine calculates the local estimator
\begin{equation}
\wmle_i = \argmax_\w \sum_{(\x,y) \in Z_i} f(y,\trans\x\w)
\end{equation}
Solving for $\wmle_i$ requires no communication with other machines.
Our goal is to provide a function $q : \Theta^m \to \Theta$ that ``merges'' these estimates.
The resulting merged estimate should have error as close to $\wmle$ as possible.

@@

The averaging estimator is defined as
\begin{equation}
\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
\end{equation}
Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal $\wmle_i-\wave$.
Then our estimator is defined as
\begin{equation}
\begin{aligned}
\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x (\hat W\alpha + \wave)\right)
\\
\wreopt &= \hat W \hat\alpha + \wave
\end{aligned}
\end{equation}
where $Z^{reopt}$ is a small dataset of size $O(m)$.
In the analysis of this estimator, we assume that $Z^{reopt}$ is independent of the $\wmle_i$;
that is, the data points in $Z^{reopt}$ were not used in the single machine optimizations.
In practice, however, this restriction is not necessary.

This estimator has the following intuitive interpretation.
We use the $\wmle_i$ estimates from the first round of communication to create a subspace $\Wave=\affspan\{\wave\}\cup{\{\wmle\}}_{i=1}^m$.
Then in the second round of optimization, we project all of our data points down into this subspace.
Because this subspace is small, we need only a small number of data points to get a good estimate of the optimum.
Because of the way we constructed $\Wave$,
the optimum in $\Wave$ will be close to the true optimum.
%The true parameter $\wstar$ should be close to this subspace.
%In the first round of optimization (determining the $\wmle_i$s),
%we're actually finding a subspace
%The estimator $\wreopt$ lives in a subs

\newpage
\section{ANALYSIS}

@@

\subsection{In the Normal Regime}
\label{sec:normal}

@@

We assume that for every estimator $\hat\w$ trained on $n$ data points,
\begin{equation}
\label{eq:assumption}
\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{I^{-1}_{\E\hat\w}}
%\hat\w \sim \normal{\E\hat\w}{\frac{1}{n}I^{-1}_{\E\hat\w}}
\end{equation}
%
%Throughout Section \ref{sec:normal} we assume that $\wave_i$ is normally distributed with mean $\wtave$ and covariance $\frac{1}{n}I^{-1}_{\wtave}$.
%\begin{equation}
%\label{eq:assumption}
%\wmle_i \sim \normal{\E\wmle_i}{\frac{1}{n}I^{-1}_{\E\wmle_i}}
%\end{equation}
%Here, $I_{\wtave}$ is the Fisher information matrix at point $\wtave$.
Here, $I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
This assumption is motivated by the central limit theorem for maximum likelihood estimation.
%There are many versions of this theorem.
Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
\begin{theorem}
\label{thm:clt}
Assume the following mild regularity conditions:
\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
\item all data points in $Z_i$ are i.i.d.;
\item the parameter space $\Theta$ is open;
\item the model is identifiable and consistent;
\item the matrix $\I_{\wtave}$ is non-singular;
\item the support set $A = \{(y,\x) : f(y,\x;\theta) > 0\}$ is independent of $\theta$;
and \item for all $(y,\x)\in A$, $f(y,\x;\theta)$ is three times differentiable with respect to $\theta$,
and the third derivative is continuous.
\end{enumerate}
Then in the limit as $n\to\infty$,
\begin{equation*}
\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\Iinv_\wstar}
\end{equation*}
Here the notation $\law$ denotes convergence in law,
which is the standard notion of convergence for a sequence of probability distributions.
\end{theorem}

@@

We provide a simple bound that shows that averaging does not improve the bias of our estimator,
but reduces the variance at the rate $O((nm)^{-1})$.
This bound is well known, but our analysis is significantly simpler and provides a sharp, easy to interpret constant factor.
The main idea of our technique is to show that the variance is normally distributed.
The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wreopt$ is the same as $\wave$, but that the bias is reduced.

>>>>

\documentclass[twoside]{article}
\usepackage{aistats2017}

@@

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{xcolor}
\usepackage[inline]{enumitem}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%\usepackage{algorithmicx}
%\usepackage{algorithm2e}

%\usepackage[authordate,bibencoding=auto,strict,backend=biber,natbib]{biblatex-chicago}
\usepackage[round]{natbib}   % omit 'round' option if you prefer square brackets
\bibliographystyle{plainnat}

\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{width=7cm,compat=1.8}

@@

\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\W}{\mathcal{\hat W}}
\newcommand{\Waff}{\mathcal{\hat A}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\wreopt}{\hat\w^{reopt}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\E\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

\newcommand{\I}{\mathcal I}
%\newcommand{\I^{-1}}{\I^{-1}}
\newcommand{\law}{\ensuremath{\xrightarrow{L}}}
\newcommand{\normal}[2]{\ensuremath{\mathcal{N}\left({{#1}},{{#2}}\right)}}
\newcommand{\trans}[1]{\ensuremath{{#1}^{\mathsf{T}}}}
\newcommand{\ltwo}[1]{{\left\lVert {#1} \right\rVert}}
\newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
\newcommand{\lzero}[1]{{\lVert {#1} \rVert}_0}
\newcommand{\proj}[1]{\pi_{{#1}}}
\newcommand{\prob}[1]{\Pr\left[{#1}\right]}

@@

\section{INTRODUCTION}

Many modern datasets are too large to fit in the memory of a single machine,
so they must be partitioned onto many machines.
%These datasets are typically partitioned onto many machines.
%Statistical learning algorithms are then run in a distributed manner.
%There are two main strategies for these distributed algorithms.
To analyze these datasets, we need distributed algorithms.
Existing distributed algorithms can be classified as either interactive or non-interactive.
In this paper we propose an algorithm that exhibits the benefits of both types.

\emph{Interactive algorithms} require many rounds of communication between machines.
These algorithms often resemble standard iterative algorithms where each iteration is followed by a communication step.
The appeal of interactive algorithms is that they enjoy the same statistical regret bounds as standard sequential algorithms.
Unfortunately, these algorithms can be too slow in practice because communication is the main bottleneck in modern distributed architectures,
and these algorithms require many rounds of communication.

\emph{Non-interactive} algorithms require only a single round of communication.
They are significantly faster than interactive algorithms
but have worse regret bounds.
Recent work (discussed in Section \ref{sec:bounds}) has shown that no non-interactive algorithm can achieve regret bounds comparable to an interactive one.
The popular large scale learning system Vowpal Wabbit cleverly uses both types of algorithms to achieve good real world performance.
A non-interactive algorithm is run first to provide a good warm start to a subsequent interactive algorithm \citep{vw,agarwal2014reliable}.
%Synonyms for non-interactive algorithms used in the literature include \emph{one-shot distributed learning} and \emph{communication efficient learning}.

In this paper, we propose a \emph{semi-interactive} distributed algorithm called XXX.
Our algorithm performs two rounds of communication,
so it is not subject to the existing regret bounds of non-interactive algorithms.
The amount of information communicated in the second round is small, however,
so our algorithm retains the speed advantages of non-interactive algorithms.

%In the remainder of this paper, we first formally describe our problem and proposed solution.
%We then provide a detailed comparison between our method and existing methods.
%We then provide a theoretical analysis.
%We conclude with an empirical analysis.

%\section{PROBLEM AND SOLUTION}

We now formally describe our problem and proposed solution.

@@

Let $\Y\subseteq\mathbb{R}$ be the space of response variables,
$X\subseteq\mathbb{R}^d$ be the space of the covariates,
and $\Theta\subseteq\mathbb{R}^d$ be the parameter space.
We assume a linear model.
The log-likelihood of data point $(\x,y)\in\X\times\Y$ given the model's true parameter $\wstar\in\Theta$ is denoted by $f(y,\trans\x\wstar)$.
Our analysis in Section \ref{sec:anal} places very mild restrictions on $f$.
In particular, $f$ need not be convex.
%For our experiments in Section \ref{sec:exp}, however, we restrict ourselves to the likelihood for logistic regression
%\begin{equation}
%f(y,\trans\x\w) =
%\end{equation}
Let $Z\subset\X\times\Y$ be a dataset of $mn$ i.i.d. observations.
Then the maximum likelihood estimator is
\begin{equation}
%\wmle=\argmax_\w \sum_{i=1}^{nm} g(y_i-\trans\x_i\w)
\wmle=\argmax_\w \sum_{(\x,y)\in Z} f(y,\trans\x\w)
\end{equation}
Assume that $Z$ has been partitioned onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
Then each machine calculates the local estimator
\begin{equation}
\wmle_i = \argmax_\w \sum_{(\x,y) \in Z_i} f(y,\trans\x\w)
\end{equation}
Solving for $\wmle_i$ requires no communication with other machines.
Our goal is to provide a function $q : \Theta^m \to \Theta$ that ``merges'' these estimates.
The resulting merged estimate should have error as close to $\wmle$ as possible.

@@

%The averaging estimator is defined as
%\begin{equation}
%\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
%\end{equation}
%Then our estimator is defined as
%\begin{equation}
%\begin{aligned}
%\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x (\hat W\alpha + \wave)\right)
%\\
%\wreopt &= \hat W \hat\alpha + \wave
%\end{aligned}
%\end{equation}
%where $Z^{reopt}$ is a small dataset of size $O(m)$.
Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal to $\wmle_i$.
Let $n^{reopt}$ be an integer less than $n$,
and $Z^{reopt}_i$ be a subset of $Z_i$ randomly sampled without replacement.
Then our estimator is defined as
\begin{equation}
\begin{aligned}
%Z_i^{proj} &= \{(y,\x\hat W) : (y,\x) \in Z_i^{reopt}\}
%\\
\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x \hat W\alpha \right)
\\
\wreopt &= \hat W \hat\alpha
\end{aligned}
\end{equation}
%In the analysis of this estimator, we assume that $Z^{reopt}$ is independent of the $\wmle_i$;
%that is, the data points in $Z^{reopt}$ were not used in the single machine optimizations.
%In practice, however, this restriction is not necessary.

%Notice that this estimator requires two rounds of communication.
There are two ways to implement this estimator.
The simplest is with two rounds of communication.
In the first round, each machine broadcasts its value of $\wmle_i$ to every other machine.
In the second round, each machine sends $Z_i^{proj}$ to a predesignated machine.
This predesignated machine performs the second optimization and outputs the final answer.
The full procedure is shown in Algorithm \ref{alg:distributed} above.
There is an alternative implementation if there is a common set of data points $Z^{common}$ possessed by each machine in addition to $Z_i$.

Our estimator has the following intuitive interpretation.
Let $\W$ be the column space of $\hat W$.
We use the $\wmle_i$ estimates from the first round of communication to create a subspace $\Wave=\affspan\{\wave\}\cup{\{\wmle\}}_{i=1}^m$.
Then in the second round of optimization, we project all of our data points down into this subspace.
Because this subspace is small, we need only a small number of data points to get a good estimate of the optimum.
Because of the way we constructed $\Wave$,
the optimum in $\Wave$ will be close to the true optimum.
%The true parameter $\wstar$ should be close to this subspace.
%In the first round of optimization (determining the $\wmle_i$s),
%we're actually finding a subspace
%The estimator $\wreopt$ lives in a subs

\begin{figure}
\hspace{-0.1in}
\begin{tikzpicture}
    [
    dot/.style = {minimum width=0.15cm,inner sep=0pt,line width=0pt,fill,circle,black,font=\small}
    ]
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.5,-0.5) (-1.05,1) (-0.1,-0.1) (0.75,-0.5) (0.45,-0.7) };
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.85,-0.65) (-1.3,1.55) (0.25,0) (2.5,-0.5) (1,-0.95)};
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-1.15,-1) (-1.5,2) (1,0) (5,0) (3,-0.95)};

\draw[thick] (-2.2,2.15) -- (5.5,1);

\node[dot] (wstar) at (-0.25,-0.25) {};
\node at (0,-0.45) {$\wstar$};

%\node[dot] (wstarproj) at (0.1,1.8) {};
%\draw (wstar) -- (wstarproj);
%\draw (0.07,1.65) -- (0.23,1.62) -- (0.26,1.78);

%\node[dot] (wavestar) at (4,0.5) {};
%\node                 at (4.5,0.3) {$\E\wave$};
\node[dot] (wave) at (3.8,1.25) {};
\node at (4.0,1.6) {$\wave$};
\node[dot] (wave) at (2.9,1.40) {};
\node at (3.1,1.75) {$\wmle_1$};
\node[dot] (wave) at (4.7,1.10) {};
\node at (4.9,1.45) {$\wmle_2$};
%\node[dot] (wave) at (3.8,1.25) {};

\node[dot] at (-1.35,2.03) {};
\node at (-1.1,2.4) {$\wreopt$};

\node at (4,-1.25) {$f(y,\trans\x\w)$};
\node at (5.75,1) {$\W$};

\end{tikzpicture}
\caption{
    Our method performs a second round of optimization to find the best parameter vector in $\W$.
    Since $\wave\in\W$, our estimator $\wreopt$ is guaranteed to have higher true likelihood with high probability.
}
\label{fig:contour}
\end{figure}

\begin{algorithm}[t]
\caption{Estimation in two rounds}
\label{alg:distributed}
\begin{algorithmic}
\State Each machine $i$ independently
\State ~~~~~loads dataset $Z_i$
\State ~~~~~calculates $\wmle_i$
\State All machines broadcast $\wmle_i$ to all other machines
\State Each machine $i$ independently
\State ~~~~~constructs $\hat W$
\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
\State ~~~~~calculates $\trans\x\hat W$
\State All machines broadcast $Z^{reopt}_i$ to a master machine
\State The master calculates $\wreopt$
\end{algorithmic}
%\label{fig:alg}
\end{algorithm}

\begin{algorithm}[t]
\caption{Estimation with common data}
\label{alg:distributed}
\begin{algorithmic}
\State Each machine $i$ independently
\State ~~~~~loads dataset $Z_i$
\State ~~~~~calculates $\wmle_i$
\State All machines broadcast $\wmle_i$ to all other machines
\State Each machine $i$ independently
\State ~~~~~constructs $\hat W$
\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
\State ~~~~~calculates $\trans\x\hat W$
\State All machines broadcast $Z^{reopt}_i$ to a master machine
\State The master calculates $\wreopt$
\end{algorithmic}
%\label{fig:alg}
\end{algorithm}

\subsection{Alternative estimators}
\label{sec:alt}
The simplest and most popular communication-efficient estimator is the averaging estimator
\begin{equation}
\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
\end{equation}

Previous analysis of $\wave$ makes a number of limiting assumptions.
\cite{mcdonald2009efficient} gave the first analysis of this estimator in the special case of L2 regularized maximum entropy models.
They provide tail bounds on $\ltwo{\wstar-\wave}$, showing that the variance of the estimator reduces as $O((nm)^{-1/2})$ with high probability,
but they could not show a reduction in bias.
Their analysis uses a martingale technique that requires the radius of the dataset be independent of the size of the dataset.
This is a particularly limiting assumption as even the simple case of Gaussian distributed data does not satisfy it.
\cite{zhang2012communication} provide a more general analysis showing that the Mean Squared Error (MSE) $\E\ltwo{\wstar-\wave}^2$ decays as $O((nm)^{-1} + n^{-2})$.
This matches the optimal MSE of $\wmle$ whenever $m<n$.
Their analysis still requires a number of technical assumptions.
For example, they assume the parameter space $\Theta$ is bounded.
This assumption does not hold under the standard Bayesian interpretation of L2 regularization as a Gaussian prior of the parameter space.
They further make strong convexity and 8\emph{th} order smoothness assumptions which guarantee that $\wmle_i$ is a ``nearly unbiased estimator'' of $\wstar$.
(As we will see in the next section,
recently developed information theoretic bounds show these assumptions are required to show a reduction in MSE to the single machine rate.)
Surprisingly, \cite{zhang2013divide} showed that in the special case of kernel ridge regression, a reduction in bias was not needed to have the MSE decay at the optimal sequential rate.
By a careful choice of regularization parameter, they caused $\wmle_i$ to have lower bias but higher variance, so that the final estimate of $\wave$ had both reduced bias and variance.

Other research has focused on modifications to the $\wave$ estimator to reduce bias.
\cite{zinkevich2010parallelized} showed that if the training sets partially overlap each other (instead of being disjoint), then the resulting estimator will have lower bias.
The authors admit their analysis is ``highly technical.''
\cite{zhang2012communication} also provided a bootstrap average estimator,
which works as follows.
Let $r\in(0,1)$, and $X_i^r$ be a bootstrap sample of $X_i$ with $rn$ data points.
Then the bootstrap average estimator is
\begin{equation}
\begin{aligned}
\wmler_i &= \argmax_\w \sum_{\x\in X_i^r} f(x;\w)
\\
\waver &= \frac{1}{m}\sum_{i=1}^m \wmler_i
\\
\wboot & = \frac{\wave-r\waver}{1-r}
\end{aligned}
\end{equation}
The intuition behind this estimator is to use the bootstrap sample to directly estimate and correct for the bias.
This estimator enjoys a MSE decaying as $O((nm)^{-1}+n^{-3})$.

\cite{liu2014distributed} propose a more Bayesian approach inspired by \cite{merugu2003privacy}.
Instead of averaging the model's parameters,
they directly ``average the models'' with the following KL-average estimator.
\begin{equation}
\wkl = \argmin_\w \sum_{i=1}^m \kl[\bigg]{p(\cdot;\wmle_i)}{p(\cdot;\w)}
\end{equation}
The minimization is performed via a bootstrap sample from the smaller models.
This method has two main advantages.
First, it is robust to reparameterizations of the model.
Second, it is statistically optimal for the class of non-interactive optimization methods.
(We show in the next section that their bound does not apply to our $\wreopt$ estimator due to our semi-interactive setting.)
The main downside of the KL-average is that the minimization has a prohibitively high computational cost.
Let $n^{kl}$ be the size of the bootstrap sample.
Then the original implementation's MSE shrinks as $O((nm)^{-1}+(nn^{kl})^{-1})$.
This implies that the bootstrap procedure requires as many samples as the original problem to get a MSE that shrinks at the same rate as the averaging estimator.
\cite{han2016bootstrap} show a method to reduce this rate to $O((nm)^{-1}+(n^2n^{kl})^{-1})$ using control variates, but the procedure remains prohibitively expensive.
Their experiments show the procedure scaling only to datasets of size $nm\approx10^4$,
whereas our experiments involve a dataset of size $nm\approx10^8$

\subsection{Performance bounds}
\label{sec:bounds}

Performance bounds come in two flavors: statistical and information theoretic.
On the statistical side, \cite{liu2014distributed} showed that for an arbitrary function $q$ \emph{which does not depend on the data},
$\ltwo{\wq-\wmle}^2$ decays as $\Omega(\gamma^2_\wstar \I^{-1}_\wstar/n^2)$.
Here $\gamma_\wstar$ is the statistical curvature of the model and $\I_\wstar$ is the Fisher information.
Furthermore, they show that $\wkl$ matches this bound.
This bound is not relevant for our estimator because our merge function depends on the data.
Furthermore, we directly optimize $\ltwo{\wq-\wstar}^2$ which is a more useful quantity in practice.

\section{ANALYSIS}
\label{sec:anal}

@@

\subsection{In the Normal Regime}
%\label{sec:normal}
%
%We say that a random vector $X$ is \emph{sub-Gaussian} with mean $\mu$ and covariance $\Sigma$
%if there exists a constant $c>0$ such that
%\begin{equation}
%\prob{\ltwo{\Sigma(X-\mu)} > t} \le c \exp(-t^2)
%\end{equation}
%We denote this relationship in symbols as $X \lesssim \normal{\mu}{\Sigma}$.
%
%\begin{theorem}
%We have that with probability at least
%\begin{equation}
%\wmle-\wstar \lesssim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
%\end{equation}
%In the limit as $n\to\infty$, we have the stronger statement that
%\begin{equation}
%\wmle-\wstar \sim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
%\end{equation}
%\end{theorem}

@@

We assume that for every estimator $\hat\w$ trained on $n$ data points,
\begin{equation}
\label{eq:assumption}
\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{\I^{-1}_{\E\hat\w}}
%\hat\w \sim \normal{\E\hat\w}{\frac{1}{n}I^{-1}_{\E\hat\w}}
\end{equation}
%
%Throughout Section \ref{sec:normal} we assume that $\wave_i$ is normally distributed with mean $\wtave$ and covariance $\frac{1}{n}I^{-1}_{\wtave}$.
%\begin{equation}
%\label{eq:assumption}
%\wmle_i \sim \normal{\E\wmle_i}{\frac{1}{n}I^{-1}_{\E\wmle_i}}
%\end{equation}
%Here, $I_{\wtave}$ is the Fisher information matrix at point $\wtave$.
Here, $\I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
This assumption is motivated by the central limit theorem for maximum likelihood estimation.
%There are many versions of this theorem.
Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
\begin{theorem}
\label{thm:clt}
Assume the following mild regularity conditions:
\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
\item all data points in $Z_i$ are i.i.d.;
\item the parameter space $\Theta$ is open;
\item the model is identifiable and consistent;
\item the matrix $\I_{\wtave}$ is non-singular;
\item the support set $A = \{(y,\x) : f(y,\x;\theta) > 0\}$ is independent of $\theta$;
and \item for all $(y,\x)\in A$, $f(y,\x;\theta)$ is three times differentiable with respect to $\theta$,
and the third derivative is continuous.
\end{enumerate}
Then in the limit as $n\to\infty$,
\begin{equation*}
\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
\end{equation*}
Here the notation $\law$ denotes convergence in law,
which is the standard notion of convergence for a sequence of probability distributions.
\end{theorem}

@@

We provide a simple bound that shows that averaging does not improve the bias of our estimator,
but reduces the variance at the rate $O((nm)^{-1})$.
Similar bounds are well known (see Section \ref{sec:alt}).
Our analysis has the following advantages:
it requires fewer assumptions, has a simpler proof, and has an easy to interpret constant factor.
The main idea of our technique is to show that the variance is normally distributed.
The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wreopt$ is the same as $\wave$, but that the bias is reduced.

\ignore{
\begin{lemma}
We have that
\begin{equation}
\E\ltwo{\wstar-\wave}^2
\le
\E\ltwo{\wstar-\E\wmle_i}^2
+
\frac{1}{m}\E\ltwo{\E\wmle_i-\wmle_i}^2
\end{equation}
\end{lemma}
\begin{proof}
It is well known that the MSE of an estimator is the sum of its bias and variance.
That is,
\begin{equation}
\E\ltwo{\wstar-\wave}^2
=
\E\ltwo{\wstar-\E\wave}^2
+
\E\ltwo{\E\wave-\wave}^2
\end{equation}
First we consider the bias.
By the linearity of expectation, we have that
\begin{align}
\E\wave
&=
\E\frac{1}{m}\sum_{i=1}^m\wmle_i
%\\&=
=
\frac{1}{m}\sum_{i=1}^m\E\wmle_i
=
\E\wmle_i
\label{eq:expwave}
\end{align}
and so the bias
$
\E\ltwo{\wstar-\E\wave}^2
=
\E\ltwo{\wstar-\E\wmle_i}^2
$.
In words, the bias of the averaging estimator is the same as the bias of the estimator on an individual machine.

Next we consider the variance.
We use the definition of $\wave$ and Equation \ref{eq:expwave} above to conclude that
\begin{align}
\wave-\E\wave
&=
\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
\\&=
\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wave\right)
\\&=
\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
%\\&\sim
%\frac{1}{m}\sum_{i=1}^m\left(\frac{1}{\sqrt n}\normal\zero{I^{-1}_{\E\wave}}\right)
%\\&=
%\frac{1}{\sqrt {nm}}\normal\zero{I^{-1}_{\E\wave}}
\end{align}
Taking the norm and squaring both sides gives
\begin{equation}
\ltwo{\wave-\E\wave}
=
\ltwo{
\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
}
\end{equation}
\end{proof}
}

\begin{theorem}
Let $t>0$.
Then with probability at least $1-\exp(-t)$,
\begin{equation}
\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wmle_i} + \sqrt\frac{v_t}{mn}
\end{equation}
where
\begin{equation}
v_t =
\tr{\I^{-1}_{\wtave}}
+ 2\sqrt{\tr \left({\I^{-2}_{\wtave}}\right)t}
+ 2\ltwo{\I^{-1}_{\wtave}}t
\end{equation}
\end{theorem}
\begin{proof}
We have by the triangle inequality that
\begin{equation}
\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wave} + \ltwo{\E\wave-\wave}
\label{eq:biasvar}
\end{equation}
The left term above captures the estimator's bias and the right term the variance.
First we consider the bias.
By the linearity of expectation, we have that
\begin{align}
\E\wave
&=
\E\frac{1}{m}\sum_{i=1}^m\wmle_i
%\\&=
=
\frac{1}{m}\sum_{i=1}^m\E\wmle_i
=
\E\wmle_i
\label{eq:expwave}
\end{align}
and so the bias term
$\ltwo{\wstar-\E\wave}
=
\ltwo{\wstar-\E\wmle_i}
$.
In words, the bias of the combined averaging estimator is the same as the bias of the estimator on an individual machine.
